Thread Pool in GeckoDB/BOLSTER - Implementing a Thread Pool for a High Performance Database System \\

### 2018-07-06 Presentations/Paper/paper_thread_pool.tex
Add figure tag to slot search code

Since $openTaskCount$ has to be altered from different threads the required decrements are ensured to be atomic. This is in opposition to the usual goal of slot-maps to improve cache locality. Atomics on x86 usually prevent their whole cache line from being used while they are changed. Thus a sparse distribution of slots should reduces collisions. To quantify this difference the baseline implementation searches for free slots linearly from the beginning, while an alternative approach tries to distribute the slots over multiple cache lines:
\begin{lstlisting}[style=CStyle]
size_t ind = 0;
// linear search, compact slot usage
for(; ind < pool->task_state_capacity && pool->task_group_states[ind].task_count; ++ind);
// distributed over different cache lines
for(; pool->task_group_states[ind].task_count; ind = (ind + 8) % MAX_NUM_TASKS);
\end{lstlisting}

@@

@@

The selection strategy of wait-slots has only a small impact on occupation as seen in \ref{fig8}. The uniform distributions advantage amounts to $0.1\%$ across different numbers of threads and is far overshadowed by differences in the threads to tasks ratio.

Figure \ref{fig9} shows the lock-free priority queue performing better. With a difference in utilization of $2\%$ when using only two threads the improvement over our mutex based implementation is not insignificant. A total occupation of over $99.5\%$ also demonstrates that most overhead in our implementation comes from the priority queue when supplying a balanced stream of tasks. With an increase in available threads the performance of the mutex solution further decreases, while the lock-free queue-variants utilization remains nearly constant.

>>>>

Since $openTaskCount$ has to be altered from different threads the required decrements are ensured to be atomic. This is in opposition to the usual goal of slot-maps to improve cache locality. Atomics on x86 usually prevent their whole cache line from being used while they are changed. Thus a sparse distribution of slots should reduces collisions. To quantify this difference the baseline implementation searches for free slots linearly from the beginning, while an alternative approach tries to distribute the slots over multiple cache lines. In Figure \ref{fig10}, we show implementations for both search approaches.

@@

\begin{figure}
	\begin{lstlisting}[style=CStyle]
	size_t ind = 0;
	// linear search, compact slot usage
	for(; ind < pool->task_state_capacity && pool->task_group_states[ind].task_count; ++ind);
	// distributed over different cache lines
	for(; pool->task_group_states[ind].task_count; ind = (ind + 8) % MAX_NUM_TASKS);
	\end{lstlisting}
	\caption{Different implementations for slot searching}
	\label{fig10}
\end{figure}

@@

The selection strategy of wait-slots has only a small impact on occupation, as seen in Figure \ref{fig8}. The uniform distributions advantage amounts to $0.1\%$ across different numbers of threads and is far overshadowed by differences in the threads to tasks ratio.

In Figure \ref{fig9}, we show that the lock-free priority queue performs better. With a difference in utilization of $2\%$ when using only two threads the improvement over our mutex based implementation is not insignificant. A total occupation of over $99.5\%$ also demonstrates that most overhead in our implementation comes from the priority queue when supplying a balanced stream of tasks. With an increase in available threads the performance of the mutex solution further decreases, while the lock-free queue-variants utilization remains nearly constant.
